---
title: "EDS 221 day 3 tasks part 2"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
library(janitor)
```

## Part 2. Real Data

```{r}
## read in data
toolik_biochem <- read_csv(here("data", "2011_Kling_Akchem.csv"), na = ".") %>% 
  janitor::clean_names()
```

Create a subset of the data that contains only observations from the \"Toolik Inlet\" site, and that only contains the variables (columns) for pH, dissolved organic carbon (DOC), and total dissolved nitrogen (TDN) (**hint:** see `dplyr::select()`). Store this subset as `inlet_biochem`. Make sure to look at the subset you\'ve created.

```{r}
inlet_biochem <- toolik_biochem %>%
  filter(site == "Toolik Inlet") %>% 
  dplyr::select(p_h, doc_u_m, tdn_u_m) 

```

Find the mean value of each column in `inlet_biochem` 2 different ways:

1.  Write a for loop from scratch to calculate the mean for each

    ```{r}
    for (i in seq(inlet_biochem)) {
      mean_val <- mean(inlet_biochem[[i]], na.rm = TRUE)
      print(mean_val)
    }
    ```

2.  Use *one other method* (e.g. `apply`, `across`, or `purrr::map_df`) to find the mean for each column.

    ```{r}
    # using across
    inlet_biochem %>% 
      summarize(across(everything(), mean, na.rm = TRUE))

    # apply
    apply(inlet_biochem, 2, mean, na.rm = TRUE)
    ```
